Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
shrinking section 1 2 3

%In the theory of Radial Basis Functions Finite-Difference (RBFFD),
%derivatives of a function $f(\rvec)$ at node $i$ are expressed as a
%linear combination of the function values at the stencil center and
%the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}). Thus
%$y$ is a discrete derivative of the vector $x$.

@@

\begin{figure}[tbh]
  \centering
  \includegraphics[width=\linewidth]{figures/rbf_stencils.pdf}
  \caption{RBFFD Stencils. Each node of the stencil is connected to
    $n_z-1$ stencil nodes in addition to itself. In the figure, node
    $A$ is connected to $B$, but $B$ is {\em not\/} connected to
    $A$. Thus adjacency graph of $A$ is not-symmetric.}
  \label{fig:rbf_stencils}
\end{figure}

@@

\begin{figure}[bth]
  \begin{center}
    \scalebox{.9}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notation relative to the matrices (upper section) and to the benchmarks (lower section.)}
  \label{tab:not}
\end{figure}

@@

Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
into actual performance. This figure provides projected best and worst
Gflop/s achievable assuming the computation is memory-bound and the
architecture reaches 150 Gflop/s. Notice that \cite{Saule13-ARXIV}
showed a higher peak bandwidth, but we will show in
Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
might achieve in these kernels. In single precision, the worse that
one could achieve by using 4 vectors and 4 matrices is much higher
than the best achievable using a classical SpMV computation. The best
performance reachable is 210Gflop/s: almost six times higher than the
peak of the classical SpMV case.

\begin{figure*}
  \centering 
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}}
%
  \subfigure[Relative performance to using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the maximum achievable performance 
    using a device with a memory to computational units of 150GB/s
    when varying the number of vectors and matrices. \todo{we could
      remove the speedup plot if we need space.}}
  \label{fig:perf_predict}
\end{figure*}

>>>>

@@

\begin{figure}[tbh]
  \centering
  \includegraphics[width=.8\linewidth]{figures/rbf_stencils.pdf}
  \caption{RBFFD Stencils. Each node of the stencil is connected to
    $n_z-1$ stencil nodes in addition to itself. In the figure, node
    $A$ is connected to $B$, but $B$ is {\em not\/} connected to
    $A$. Thus adjacency graph of $A$ is not-symmetric.}
  \label{fig:rbf_stencils}
\end{figure}

@@

\begin{figure}[bth]
  \begin{center}
    \scalebox{.8}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notation relative to the matrices (upper section) and to the benchmarks (lower section.)}
  \label{tab:not}
\end{figure}

@@

Figure~\ref{fig:perf_predict} shows how these ratios translate
into actual performance. This figure provides projected best and worst
Gflop/s achievable assuming the computation is memory-bound and the
architecture reaches 150 Gflop/s. Notice that \cite{Saule13-ARXIV}
showed a higher peak bandwidth, but we will show in
Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
might achieve in these kernels. In single precision, the worse that
one could achieve by using 4 vectors and 4 matrices is much higher
than the best achievable using a classical SpMV computation. The best
performance reachable is 210Gflop/s: almost six times higher than the
peak of the classical SpMV case.

\begin{figure}
  \centering 
  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
  \caption{Estimation of the maximum achievable performance using a
    device with a memory to computational units of 150GB/s when
    varying the number of vectors and matrices.}
  \label{fig:perf_predict}
\end{figure}
